Relevance-based Quantization of Scattering Features for Unsupervised Mining of Environmental Audio

### 2017-09-12 paper/structureScene.tex
rewrite of intro step 1

This works falls in the realm of ecoacoustics, an emerging field in computational bioacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause, wimmer2013sampling}. %Despite identified needs, the field of signal-based ecoacoustics is still in its infancy.
%Consequently, few well-designed datasets are readily available for evaluation purposes.

An important problem in this field is the characterization of auditory scenes for the purposes of similarity estimation.

To this aim, one needs to design an algorithm to perform an acoustic scene similarity retrieval, where the scenes judged ``most similar'' to a given query scene are extracted from some larger dataset. This requires the auditory scenes to be represented in some way that captures their distinctive qualities. This has been achieved by bag-of-frames approaches \cite{aucouturier2007bag}, which attempts to describe the auditory scene recording using summary statistics. Unfortunately, since this only captures the average structure of the scene, the approach often fails when presented scenes which are dynamic and contain episodic, distinct sound events that discriminate the scene from others. Indeed, experiments in cognitive psychology \cite{dubois2006cognitive} and cognitive neurosciences \cite{nelken2004processing} suggest that this matches more closely the way that humans perceive acoustic scenes. We believe that the failure to model such distinct events is one of the reason why this approach is found to be not sufficient when considering \cite{}.

To design better representations for the similarity retrieval task, we draw inspiration from a closely related, and more mature, field of investigation: the classification of acoustic scenes. This task, consisting of predicting scene labels from audio recordings, is narrower in its range of applications, but has the advantage of being rooted in the cognitive psychology of categorization \cite{dubois2006cognitive,guastavino_ideal_2006}. In addition, it has received significant attention from the data mining community for some time, resulting in the availability of well-designed datasets and competitive algorithms for solving it.

@@

For the task at at hand, the scattering features have the valuable advantage of not requiring training and thus potentially lead to wider range of application than trained. Indeed, the task at hand is the mining of yet unheard audio datasets and the properties of the audio that are of interest remains to be

The next question is the choice of a classifier. Here, a class of particularly successful algorithms train a classifier, such as a support vector machine (SVM) \cite{cristianini2000introduction}, to discriminate between short extracts of an auditory scene, and at the classification stage divide the signal into short extracts which are classified separately before pooling the classification results to reach a final decision. This last step is known as late integration and is crucial to the success of these methods, since it allows the classifier to concentrate on the short, distinct sound events that discriminate between scenes. The success of the late integration technique again leads credence to the idea that rare, distinct events in an auditory scene are sufficient to characterize it almost completely. Combining scattering transforms for low-level representation with late-integrated SVMs for classifying the scene at a higher level, we obtain a competitive auditory scene classification system, demonstrating the usefulness of both concepts for this task.

Returning to scene retrieval setting, we cannot use supervised models such as SVMs, but we can emulate their selection behavior by quantizing the auditory scene and defining a distance over this quantization. We therefore propose a new model for acoustic scenes, where the signal is represented at sub-second scales by scattering transforms, while larger scales are captured by an unsupervised method that quantizes the scattering coefficients into a given number of clusters. We then use these clusters to define a set of distances that can subsequently be used for similarity retrieval. Testing this approach on an scene retrieval task, we obtain significant improvements over traditional bag-of-frames and summary statistics models, applied both to MFCCs and scattering coefficients.

Motivations of the proposed approach and a brief review of the state of the art in the field of acoustic scene modeling is given in Section \ref{sec:soa}. We describe the scattering transform in Section \ref{sec:scattering}, discuss processing of the resulting features in Section \ref{sec:design} and propose a cluster-based scene description in Section \ref{sec:object}. In Section \ref{sec:experiments}, several experiments are described for the acoustic scene classification task and the acoustic scene similarity retrieval task. Results are subsequently reported in Section \ref{sec:results}.

@@

\section{Supervised classification system}
\label{sec:supervised}

We are now ready to present the supervised method that we will apply to the acoustic scene classification task using scattering transforms. It consists of two parts, a classifier and a temporal integration method. While the classifier is an SVM, we shall consider two integration approaches, early and late.

\subsection{Support vector machine classifier}
\label{sec:svm}

A support vector machine (SVM) is a binary discriminative classifier that attempts to find a separating hyperplane between a pair of classes of that provides the maximal margin between the two sets \cite{Cortes1995}. From another point of view, it attempts to find a direction on which to project in order to maximize separation between two point clouds. Since it is not a generative, but discriminative model, it is able to pick out those feature vectors that best characterize a particular class of auditory scenes compared to another. This property, coupled with its ease of use and efficient implementations \cite{libsvm} have made SVMs a popular choice for many classification systems.

At the classification stage, the SVM gives a class label as output, which is one of the two classes it models. For multi-class settings, a one-\vs-one approach is often used, where an SVM is trained for each pair of classes. During classification, the outputs of all pairwise SVMs is computed and the class selected the largest number of times is chosen.

Finally, the SVM can also be extended to handle nonlinear decision surfaces. This is obtained using the ``kernel trick'', which exploits the fact that all calculations in the SVM rely only on the inner product between vectors. By replacing this standard inner product by some other positive definite symmetric mapping, such as a Gaussian radial basis function (RBF), this is equivalent to performing a linear SVM after mapping the feature vectors into some high-dimensional space. A nonlinear mapping then results in a nonlinear decision surface in the original space.

\subsection{Early \vs late temporal integration}
\label{sec:eili}

Owing to the scarcity of salient events in many natural scenes,
fine-grained classification is only made
possible by integrating signal information over a long temporal context.
Indeed, whereas a few seconds is often sufficient to recognize a speaker,
a musical instrument, or a genre, up to $30$ seconds may be required
to disambiguate two classes of auditory scenes with overlapping semantic content, \eg a train from a subway station or a quiet street from a park.
Depending on whether aggregation is performed in feature space or in decision space (the output of a classifier) the corresponding method is referred to as early or late integration.

A straightforward application of early integration consists in summarizing the multivariate time series of scattering coefficients over the full duration of the auditory scene by computing their average values over time.
In the definition of the scattering transform (see Section \ref{sec:scattering}), this is equivalent to increasing the support $T$ of the low-pass filter $\boldsymbol{\phi}(t)$ up to infinity. With a slight abuse of notation, we denote the summarized features by
\begin{equation}
\mathbf{S}\boldsymbol{x}(\gamma) =
\int_{-\infty}^{+\infty} \mathbf{S}\boldsymbol{x}(t,\gamma)\;\mathrm{d}t\mbox{.}
\end{equation}

Conversely, a late integration scheme relies on probabilistic assignments $\mathbb{P}\left[y \,\vert\, \mathbf{S}\boldsymbol{x}(t,\gamma) \right]$ over short-term windows of length $T$ obtained from some supervised model, such as a classifier. These are subsequently aggregated to produce a final decision
\begin{equation}
\hat{y} = \arg \max_{y} \rho\Big(\big\{ \mathbb{P}\left[y \,\vert\, \mathbf{S}\boldsymbol{x}(t,\gamma) \right] \big\}_{t} \Big)\mbox{,}
\end{equation}
where $\hat{y}$ is the estimated class label and $\rho$ is a reduction function, such as sum, product, or majority vote \cite{Kittler1998}.

A potential drawback of early integration is that it reduces a sequence of feature vectors to a single average. If the underlying signal is static in nature, this may be advantageous, as the temporal variation of the feature vectors constitutes an unwanted source of variability. For many audio signals, and in particular auditory scenes, this is very far from the truth, and prevents us from identifying the distinct sound events that characterize the signal as these may be drowned out by the background during the averaging step. In this case, a late integration approach would perform better, since it operates on the individual feature vectors and only the classification results are aggregated over time. Using experiments in acoustic scene classification below, we will demonstrate that the late integration approach does indeed succeed over early integration when scattering coefficients are used.

@@

\begin{table}
\begin{center}
\caption{
\label{table:dcase2013}
Classification accuracies on the DCASE 2013 dataset with various settings of features, temporal integration strategies, and support vector machine kernels.
}
\begin{tabular}{llll}
             & MFCCs         & scattering & log-scattering  \\
             \hline
\emph{early}, linear  & $54\pm19$   & $62\pm\phantom{0}6$  & $66\pm11$     \\
\emph{early}, RBF     & $47\pm16$  & $61\pm\phantom{0}9$  & $63\pm\phantom{0}6$   \\
\emph{late}, linear  & $60\pm\phantom{0}9$ & $70\pm\phantom{0}8$  & $75\pm\phantom{0}5$   \\
\emph{late}, RBF     & $68\pm14$ & $73\pm\phantom{0}6$  & $\mathbf{78}\pm\phantom{0}6$   \\
\end{tabular}
\end{center}
\end{table}

\subsection{Experiment \#1: acoustic scene classification}

Our first task is to compare the performance of the standard scattering transform, the logarithmically compressed scattering (log-scattering) transform, and baseline MFCCs. Secondly, we evaluate the two temporal integration approaches presented above, \emph{early} and \emph{late}.

The different systems are evaluated on the private part of the DCASE 2013 dataset using five-fold stratified cross validation. The folds are identical to those used in the original challenge. The slack parameter $C$ of the SVM is here set to $1$. Both linear and Gaussian RBF kernels are used. For the latter, a local scaling method is applied \cite{selfTuneManor2004}. All feature vectors, that is scattering frames and MFCCs, are standardized prior to be given to the classifier. The standardization is computed with respect to the train/test splits of the cross validation scheme (see Section \ref{sec:stand}). The optimal scaling parameter $q$ of the RBF kernels is learned on the public part of the DCASE 2013 dataset, using five-fold stratified cross validation.  For these experiments, both \emph{early} and \emph{late} integration strategies are evaluated (see Section \ref{sec:eili}).

\subsection{Experiment \#2: acoustic scene similarity retrieval}

@@

The results on acoustic scene classification illustrate the superior performance of the log-scattering transform coupled with late integration, confirming the ideas motivating the construction of our cluster-based auditory scene representation. Using this representation for auditory scene similarity retrieval, we obtain results that improve significantly over traditional BOF and summary approaches, both with log-scattering and MFCCs features.

\subsection{Experiment \#1: acoustic scene classification}

%\begin{figure}
%\begin{center}
%\includegraphics[width=\columnwidth]{bw/folds.png}
%\caption{\label{fig:folds}
%Classification accuracies over the five folds of the DCASE 2013 private dataset, for various acoustic descriptors and temporal integration strategies.  Performance on each fold is depicted using a given symbol.
%The chosen classifier is a support vector machine (SVM) whose kernel is a Gaussian radial basis function (RBF).
%The mean accuracies and fold-wise standard deviations of each experiment are reported in Table \ref{table:dcase2013}.}
%\end{center}
%\end{figure}

\begin{table}
\begin{center}
\caption{Class-wise accuracies on the DCASE 2016 dataset.\label{table:dcase2016}}
\begin{tabular}{lccc}
Features & MFCCs & log-scattering & log-scattering \\
Subset & development & development & evaluation \\
\midrule
beach & $58.3 \pm 28.2$ & $83.5 \pm \phantom{0}7.1$ & $80.8$\\
bus & $58.2 \pm 25.4$ & $88.8 \pm 11.4$ & $92.3$ \\
cafe/restaurant & $77.4 \pm 14.7$ & $64.5 \pm \phantom{0}8.3$ & $50.0$ \\
car & $92.3 \pm \phantom{0}7.7$ & $94.9 \pm \phantom{0}6.1$ & $96.2$ \\
city\_center & $96.4 \pm \phantom{0}3.8$ & $91.7 \pm \phantom{0}9.4$ & $84.6$ \\
forest\_path & $67.7 \pm 30.8$ & $93.8 \pm \phantom{0}7.8$ & $96.2$ \\
grocery\_store & $44.7 \pm 24.8$ & $90.9 \pm \phantom{0}6.9$ & $84.6$ \\
home & $44.8 \pm 15.3$ & $56.2 \pm 23.9$ & $80.8$ \\
library & $54.6 \pm 20.3$ & $82.0 \pm 13.0$ & $65.4$ \\
metro\_station & $94.7 \pm \phantom{0}6.8$ & $96.0 \pm \phantom{0}2.3$ & $96.2$ \\
office & $95.7 \pm \phantom{0}7.5$ & $87.5 \pm 21.7$ & $100$ \\
park & $52.4 \pm 10.4$ & $75.4 \pm \phantom{0}7.4$ & $65.4$ \\
residential\_area & $53.2 \pm 17.0$ & $44.2 \pm 15.0$ & $69.2$ \\
train & $44.5 \pm 23.1$ & $58.3 \pm 10.0$ & $53.8$ \\
tram & $52.1 \pm 11.9$ & $83.7 \pm 12.2$ & $96.2$ \\
\bottomrule
Average & $65.8 \pm 19.3$ & $79.4 \pm 15.6$ & $80.8 \pm 16.4$ \\
\end{tabular}
\end{center}

\end{table}

We report experiments on acoustic scene classification, following the original methodology of the DCASE 2013 challenge, and extend our approach to its 2016 edition.

\subsubsection*{Role of logarithmic compression}

On the DCASE 2013 dataset, logarithmic compression provides a small, albeit consistent boost to acoustic scene classification.
Indeed, as discussed in subsection \ref{sec:logcomp}, and observed in Figure \ref{fig:compression}, the statistical assumption of normality for scattering coefficients is only approximately satisfied once mapped to a logarithmic, decibel-like scale. This transformation improves the discriminative power of the SVMs trained on scattering coefficients, regardless of the chosen temporal integration strategy (early or late) and the inner product kernel (linear or Gaussian RBF).

\subsubsection*{MFCCs \vs scattering}

%As shown in Figure \ref{fig:folds}, scattering coefficients outperform MFCCs over all but one fold of the DCASE 2013 private dataset when applying early integration, and over all folds when applying late integration.

As shown on Table \ref{table:dcase2013}, scattering coefficients outperform MFCCs on the DCASE 2013 private dataset when applying early and integration.
This is in accordance with previous results on similarity retrieval, as well as comparisons between mel-frequency coefficients and scattering coefficients for tasks ranging from musical genre recognition and phone segment classification \cite{Anden2014} to environmental sound classification \cite{Salamon2015}.

In order to evaluate the generality of our findings, we have taken part in the acoustic scene classification task of the 2016 edition of the DCASE challenge, which was evaluated on a much larger dataset compared to DCASE 2013. For this reason, training support vector machines with the Gaussian RBF kernel proved computationally too demanding and we instead used a linear kernel. Results on DCASE 2013 (see Table \ref{table:dcase2013}) suggests that this provides comparable, albeit slightly worse results. The results on DCASE 2016 are reported on Table \ref{table:dcase2016}.

We compare our results with baseline classifier using linear SVMs and MFCC features, which achieves an average accuracy of $65.5\%$. With the scattering transform, however, the average accuracy increases to $79.4\%$. This accuracy increases slightly to $80.8\%$ when calculated on the held-out evaluation dataset.\footnote{\url{http://www.cs.tut.fi/sgn/arg/dcase2016/task-results-acoustic-scene-classification}} The small change in overall accuracy and relatively stable class-wise accuracies suggest that there is little overfitting in the model, an indication of good generalization power.

\subsubsection*{Early \vs late integration}
On the DCASE 2013 dataset, late integration, \ie building a prediction by majority voting over $128$ local windows of duration $T=372\,\mathrm{ms}$, outperforms early integration, \ie classifying the acoustic scene with a global window of duration $T=24\,\mathrm{s}$.
This suggests that there is relevant variability in scales beyond the window size $T = 372~\mathrm{ms}$ that is not captured by averaging the scattering transform in time (early integration). One potential solution to this is to increase the scale of the second-order modulation wavelets from $J_2 = 12$ octaves to some larger value. This does not result in any performance gain, however (results not shown). What does help is the late integration, since it allows the SVM to successfully classify the few short extracts that characterize the scene which are then used to classify the whole scene through the majority vote. Since the scattering transform, even with larger values of $J_2$, is not able to separate out these short, discriminative sound events, delegating this task to the late-integrated results in a more powerful classification system.

\subsection{Experiment \#2: acoustic scene similarity retrieval}

>>>>

This works falls in the realm of ecoacoustics, an emerging field in computational bioacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause}. %Despite identified needs, the field of signal-based ecoacoustics is still in its infancy.
%Consequently, few well-designed datasets are readily available for evaluation purposes.

An important problem in this field is that analysing the recorded data to identify the quantities of interest is very costly \cite{wimmer2013sampling} and needs some sort of pre-screening to reduce the need for human expert listening and annotation.

The most straightforward way is to assume a closed set of events of interests like sounds of animals expected to live near the acoustic sensors, train models for them and perform automatic annotation \cite{}. By doing so, a period of time, say a day is represented as the number of events within the predefined typology that have been detected in this period of time. This approach allows the scientist to reduce drastically the amount of information that needs to be processed.  Though, this approach have two drawbacks. The first is the reliance to trained models whose behavior on unseen data (different sensors, ...) is prone to errors and can hardly be asserted. And, more importantly, it is based on \textit{a priori} knowledge, and thus cannot be considered for exploratory analysis where the quantities of interest are yet to be defined.

and annotation to whether recurring patterns  or sparsely distributed events to study anomalous or scarce behaviors. The latter allowing the scientist to then annotate only a representative audio of the periods of time with recurring patterns and the former to potentially discover unforseen phenomena.

To this aim, one needs to design an algorithm to perform an acoustic similarity retrieval, where the audio fragments judged ``most similar'' to a given query section are extracted from some larger dataset. This requires the audio fragments to be represented in some way that captures their distinctive qualities. This has been achieved by bag-of-frames approaches in the litterature\cite{aucouturier2007bag}, which attempts to describe the auditory scene recording using summary statistics. Unfortunately, since this only captures the average structure of the scene, the approach often fails when presented scenes which are dynamic and contain episodic, distinct sound events that discriminate the scene from others. Indeed, experiments in cognitive psychology \cite{dubois2006cognitive} and cognitive neurosciences \cite{nelken2004processing} suggest that this matches more closely the way that humans perceive acoustic scenes. We believe that the failure to model such distinct events is one of the reason why this approach is found to be not sufficient when considering decent size evaluation datasets \cite{lagrange:hal-01082501}.

%To design better representations for the similarity retrieval task, we draw inspiration from a closely related, and more mature, field of investigation: the classification of acoustic scenes. This task, consisting of predicting scene labels from audio recordings, is narrower in its range of applications, but has the advantage of being rooted in the cognitive psychology of categorization \cite{dubois2006cognitive,guastavino_ideal_2006}. In addition, it has received significant attention from the data mining community for some time, resulting in the availability of well-designed datasets and competitive algorithms for solving it.

@@

For the task at at hand, the scattering features have the valuable advantage of not requiring training and thus potentially lead to wider range of application than trained features. Indeed, the task at hand is the mining of yet unheard audio datasets and the properties of the audio that are of interest remains to be defined. Thus leading to an unserpervised setting.

%The next question is the choice of a classifier. Here, a class of particularly successful algorithms train a classifier, such as a support vector machine (SVM) \cite{cristianini2000introduction}, to discriminate between short extracts of an auditory scene, and at the classification stage divide the signal into short extracts which are classified separately before pooling the classification results to reach a final decision. This last step is known as late integration and is crucial to the success of these methods, since it allows the classifier to concentrate on the short, distinct sound events that discriminate between scenes. The success of the late integration technique again leads credence to the idea that rare, distinct events in an auditory scene are sufficient to characterize it almost completely. Combining scattering transforms for low-level representation with late-integrated SVMs for classifying the scene at a higher level, we obtain a competitive auditory scene classification system, demonstrating the usefulness of both concepts for this task.

%we can emulate their selection behavior by quantizing the auditory scene and defining a distance over this quantization.
We also propose a new model for acoustic scenes, where the signal is represented at sub-second scales by scattering transforms, while larger scales are captured by an unsupervised method that quantizes the scattering coefficients into a given number of clusters. We then use these clusters to define a set of distances that can subsequently be used for similarity retrieval. Testing this approach on an scene retrieval task, we obtain significant improvements over traditional bag-of-frames and summary statistics models, applied both to MFCCs and scattering coefficients.

Motivations of the proposed approach and a brief review of the state of the art in the field of acoustic scene modeling is given in Section \ref{sec:soa}. We describe the scattering transform in Section \ref{sec:scattering}, discuss processing of the resulting features in Section \ref{sec:design} and propose a cluster-based scene description in Section \ref{sec:object}. In Section \ref{sec:experiments}, several experiments are described for the acoustic scene similarity retrieval task. Results are subsequently reported in Section \ref{sec:results}.

@@

@@

@@

%The results on acoustic scene classification illustrate the superior performance of the log-scattering transform coupled with late integration, confirming the ideas motivating the construction of our cluster-based auditory scene representation. Using this representation for auditory scene similarity retrieval, we obtain results that improve significantly over traditional BOF and summary approaches, both with log-scattering and MFCCs features.
